Ultrafast manipulation of Heisenberg exchange and Dzyaloshinskii–-Moriya interactions in antiferromagnetic insulators

### 2019-02-11 Paper/paper.tex
Mod picture and add NN interaction

\begin{equation}
\label{MKMH}
\hat{H}^0 = \hat{\mathcal{H}}_t + \hat{\mathcal{H}}_{\text{SOI}} + \hat{\mathcal{H}}_{\text{int}}
\end{equation}
Where:
\begin{align}
&\hat{\mathcal{H}}_t = - \sum_{\langle i,j \rangle, \sigma} t_1\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} - \sum_{\langle \langle i,j \rangle \rangle, \sigma} t_2\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} \\
&\hat{\mathcal{H}}_{\text{SOI}} = \sum_{\langle \langle i,j \rangle \rangle, \sigma} i\Delta\nu_{ij}\sigma^z_{\sigma, \sigma}\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} \\
&\hat{\mathcal{H}}_{\text{int}} = \text{U}\hat{D}
\end{align}

Where $\hat{c}_{i \sigma}^\dagger$ ($ \hat{c}_{i \sigma}$) creates (anihilates) an electron at site $i$ in spin state $\sigma$. $\hat{D} = \sum_{i=1}^M \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}$ is the doublon number operator and $\text{U}$ is the on-site interaction. $t_1$, $t_2$ are the hopping amplitudes originating from both kinetic hopping. Where $\Delta$ is the intrinsic spin orbit coupling constant. $\nu_{ij}=\pm 1$ depending on whether the electron traversing from $i$ to $j$ makes a right ($+1$) or a left turn ($-1$). $\sigma^{z}$ is the third Pauli matrix. 
\begin{figure}[t]
\includegraphics[width=0.6\columnwidth]{../Figures/kmh.png}
\caption{A honycomb cell with NN hopping $t_1$, NNN hopping $t_2$ and intrinsic SOI $\Delta$. $\nu_{ij} = \pm 1$ depending on whether the electron traversing from $i$ to $j$ makes a right ($+1$) or a left turn ($-1$).}
\label{fig1}
\vspace*{-6pt}
\end{figure}
We will assume a half filling system in which the strength of the on-site interaction $\text{U}$ is much larger than the hopping amplitudes. In the strong coupling limit any state with a nonzero number of double occupancies ($d \neq 0$) will have much larger energy than those with $d=0$. We can therfore obtain an effective Hamiltonian acting on the $d=0$ subspace by standard second order perturbation techniques in the hopping terms. We can then obtain the effective spin Hamiltonian by using the relations:
\begin{align}
\hat{c}_{i \sigma}^\dagger \hat{c}_{i \sigma'} &= \delta_{\sigma \sigma'} \frac{1}{2} (n_{i \uparrow} + n_{i \downarrow}) + \bs{S}_i\cdot\bs{\sigma}_{\sigma', \sigma} \label{SpinOperatorInv1}\\ 
\hat{c}_{i \sigma} \hat{c}_{i \sigma'}^\dagger &= \delta_{\sigma \sigma'} \frac{1}{2} (2 - n_{i \uparrow} - n_{i \downarrow}) - \bs{S}_i\cdot\bs{\sigma}_{\sigma, \sigma'} \label{SpinOperatorInv2}
\end{align}
Following this procedure we obtain the following effective spin Hamiltonian:

@@

\begin{figure}[t]
\includegraphics[width=0.6\columnwidth]{../Figures/NNvsNNN.pdf}
\caption{For circularly polarized light we obtain $\tilde{J}_1 = \frac{J_{1,ij}}{J_{1,ij}^0}$ and $\tilde{D}_{2} = \frac{D_{2,ij}}{D_{2,ij}^0}$ as function of $\mathcal{E} = \frac{eaE_0}{\omega}$, where $e$ is the electron charge and $a$ is the lattice constant. The results are obtain for (units of $\hbar=t_1=1$) $t_2 = 0.1$, $\Delta = 0.5$, $\text{U} = 10$ and $\omega = 6, 16$. Solid lines are for $\omega = 6$ and dashed lines are for $\omega = 16$. For $\tilde{J}_1$, similar results are obtained by Mentink et Al. \cite{Mentink2015}.}
\label{fig2}
\vspace*{-6pt}
\end{figure}
A similar Hamiltonian including a NN exchange term and a NNN DMI term was first proposed by S. A. Owerre to model honeycome topological magnon insulators \cite{Owerre2016} \cite{Elyasi}. Additionally, experimental results regarding topological properties of spin waves in honeycomb ferromagnet Cr$I_3$ can only be understood by considering this Hamiltonian \cite{Chen2018}. This model is also relevant for the study of Spin Hall effects of Weyl magnons \cite{Zyuzin2018} \cite{Sekine2015}.

Now, a Hamiltonian with the form $\hat{H} = \sum_{\langle i,j \rangle} J_1 \bs{S}_i\bs{S}_j + \sum_{\langle \langle i,j \rangle \rangle} J_2\bs{S}_i\bs{S}_j$ is known as the $J_1$-$J_2$ Heisenberg model and in a 2D honeycomb lattice it exhibits N\'eel order for $J_2 < J_1 / 6$ and for $J_2 > J_1 / 6$ spin density waves (SDW) appear \cite{Mulder2010}. In the presence of DMI alone there will always be SDW in the plane perpendicular to $\bs{D}$ \cite{Uchida2006}. In Hamiltonian \ref{MKMHeff} we expect SDW to appear in the ground state and the SDW wavevector will be determined by a function of the parameters of this model. In the next section we will do a numerical of study on how modifying the ratio between NN and NNN spin interaction factors can change the spin state of the system.

>>>>

\begin{equation}
\label{MKMH}
\hat{H}^0 = \hat{\mathcal{H}}_t + \hat{\mathcal{H}}_{\text{SOI}} + \hat{\mathcal{H}}_{\text{int}}
\end{equation}
Where:
\begin{align}
&\hat{\mathcal{H}}_t = - \sum_{\langle i,j \rangle, \sigma} t_1\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} - \sum_{\langle \langle i,j \rangle \rangle, \sigma} t_2\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} \\
&\hat{\mathcal{H}}_{\text{SOI}} = \sum_{\langle \langle i,j \rangle \rangle, \sigma} i\Delta\nu_{ij}\sigma^z_{\sigma, \sigma}\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} \\
&\hat{\mathcal{H}}_{\text{int}} = \text{U}_{00}\sum_{i=1}^M \hat{n}_{i\uparrow}\hat{n}_{i\downarrow} + \frac{1}{2}\sum_{\langle i,j \rangle, \sigma \sigma'} \text{U}_{ij}\hat{n}_{i\sigma}\hat{n}_{j\sigma'}
\end{align}

Where $\hat{c}_{i \sigma}^\dagger$ ($ \hat{c}_{i \sigma}$) creates (anihilates) an electron at site $i$ in spin state $\sigma$, $\text{U}_{00}$ and $\text{U}_{ij}$ are the on-site and NN Coulomb interactions. $t_1$, $t_2$ are the hopping amplitudes originating from both kinetic hopping. $\Delta$ is the intrinsic spin orbit coupling constant. $\nu_{ij}=\pm 1$ depending on whether the electron traversing from $i$ to $j$ makes a right ($+1$) or a left turn ($-1$). $\sigma^{z}$ is the third Pauli matrix. 
In practice the NN Coulomb interaction can be effectively approximated by a reduced on-site interaction $\text{U} = \text{U}_{00} - \bar{\text{U}}$ where $\bar{\text{U}}$ is an average of the NN Coulomb interacion \cite{Schuler2013}, we therefore can take:
\begin{equation}
\hat{\mathcal{H}}_{\text{int}} \approx \text{U}\sum_{i=1}^M \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}
\end{equation}
With $\text{U} = \text{U}_{00} - \bar{\text{U}}$.

\begin{figure}[t]
\includegraphics[width=0.6\columnwidth]{../Figures/kmh.png}
\caption{A honycomb cell with NN hopping $t_1$, NNN hopping $t_2$ and intrinsic SOI $\Delta$. $\nu_{ij} = \pm 1$ depending on whether the electron traversing from $i$ to $j$ makes a right ($+1$) or a left turn ($-1$).}
\label{fig1}
\vspace*{-6pt}
\end{figure}
Let us define $\hat{D} = \sum_{i=1}^M \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}$ the doublon number operator so that $\hat{\mathcal{H}}_{\text{int}} \approx \text{U}\hat{D}$ with eigenvalues $d$ and projection operators $\hat{P}_d$. We will assume a half filling system in which the strength of the on-site interaction $\text{U}$ is much larger than the hopping amplitudes. In the strong coupling limit any state with a nonzero number of double occupancies ($d \neq 0$) will have much larger energy than those with $d=0$. We can therfore obtain an effective Hamiltonian acting on the $d=0$ subspace by standard second order perturbation techniques in the hopping terms. We can then obtain the effective spin Hamiltonian by using the relations:
\begin{align}
\hat{c}_{i \sigma}^\dagger \hat{c}_{i \sigma'} &= \delta_{\sigma \sigma'} \frac{1}{2} (n_{i \uparrow} + n_{i \downarrow}) + \bs{S}_i\cdot\bs{\sigma}_{\sigma', \sigma} \label{SpinOperatorInv1}\\ 
\hat{c}_{i \sigma} \hat{c}_{i \sigma'}^\dagger &= \delta_{\sigma \sigma'} \frac{1}{2} (2 - n_{i \uparrow} - n_{i \downarrow}) - \bs{S}_i\cdot\bs{\sigma}_{\sigma, \sigma'} \label{SpinOperatorInv2}
\end{align}
Following this procedure we obtain the following effective spin Hamiltonian:

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{../Figures/NNvsNNN.pdf}
\caption{For circularly polarized light we obtain $\tilde{J}_1 = \frac{J_{1,ij}}{J_{1,ij}^0}$ and $\tilde{D}_{2} = \frac{D_{2,ij}}{D_{2,ij}^0}$ as function of $\mathcal{E} = \frac{eaE_0}{\omega}$, where $e$ is the electron charge and $a$ is the lattice constant. The results are obtain for (units of $\hbar=t_1=1$) $t_2 = 0.1$, $\Delta = 0.5$, $\text{U} = 10$ and $\omega = 6, 16$. Solid lines are for $\omega = 6$ and dashed lines are for $\omega = 16$. For $\tilde{J}_1$, similar results are obtained by Mentink et Al. \cite{Mentink2015}.}
\label{fig2}
\vspace*{-6pt}
\end{figure}
A similar Hamiltonian including a NN exchange term and a NNN DMI term was first proposed by S. A. Owerre to model honeycome topological magnon insulators \cite{Owerre2016} \cite{Elyasi}. Additionally, experimental results regarding topological properties of spin waves in honeycomb ferromagnet Cr$I_3$ can only be understood by considering this Hamiltonian \cite{Chen2018}. This model is also relevant for the study of Spin Hall effects of Weyl magnons \cite{Zyuzin2018} \cite{Sekine2015}.

Now, a Hamiltonian with the form $\hat{H} = \sum_{\langle i,j \rangle} J_1 \bs{S}_i\bs{S}_j + \sum_{\langle \langle i,j \rangle \rangle} J_2\bs{S}_i\bs{S}_j$ is known as the $J_1$-$J_2$ Heisenberg model and in a 2D honeycomb lattice it exhibits N\'eel order for $J_2 < J_1 / 6$ and for $J_2 > J_1 / 6$ spin density waves (SDW) appear \cite{Mulder2010}. In the presence of DMI alone there will always be SDW in the plane perpendicular to $\bs{D}$ \cite{Uchida2006}. In Hamiltonian \ref{MKMHeff0} we expect SDW to appear in the ground state and the SDW wavevector will be determined by a function of the parameters of this model. We have shown that we can modulate these parameters thus providing a techique to manipulate the SDW in the ground state.
